---
title: "EcoTopic2_eBird"
author: "Stellah"
date: "2026-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Research Question: 
How is noise pollution related to population distribution of birds across NY state? 

Hypotheses: 

Increased proximity to noise pollution will be associated with a decreased bird population as noise pollution can disrupt breeding and communication of birds as well as increase stress levels. 

Increased volume of noise pollution will be associated with a decreased bird population as higher volumes of noise pollution are more disruptful to breeding and communication of birds as well as increasing stress. 


```{r}
#general definitions:
##NP = noise pollution 
```

Library
```{r}
#library any new packages we need to analyze 
#comment what the package will be used for 
```

Loading Data
```{r}
#load in data 
##NP -- DOT Transportation 2020, clipping NYS
##EBird -- NY State data for 2020
##Counties -- NY State country boundaries

##naming convention is all lowercase with underscores and use "data" at end for raw data 

#head() all of these datasets to get a feel for their contents 
```

NP Histogram
```{r}
#make a histogram of the distribution of decibels of NP in NY to understand the data

#are there clear bin lines that we can create to create an index for NP? Then could use that throughout our analysis to simplify the NP data -- if yes, explore with boxplots
```

NP aggregation by county & mapping
```{r}
#aggregate NP by county in NY 

#map decibels of NP across NY by county 
```

Bird aggregation by county investigation
```{r}
#basic map of all bird sightings in NY in 2020 to see if aggregating by county (which we suspect will be effective) is actually a good way to group the data

#aggregate bird sightings by county 

#make a histogram of the distribution of bird observations by county

#make a bivariate map of the bird count and NP by county 
```

Exploring overall linear relationship
```{r}
#create a scatterplot for the relationship between noise pollution and bird sightings (expecting a negative linear relationship)

#run a lm on this relationship 
```

Diving into specific species  
```{r}
#find the top 5 most common birds in the dataset and compare them 

#create boxplots for the relationship between noise pollution index and each species count (differentiate by color)
```

NYC Case Study 
```{r}
##make a histogram of the distribution of decibels of NP in NYC to understand the data

###may need to create a new index for this -- as NYC will be louder than the rest of NY

#map all bird sightings to familiarize with distribution across the city  
###maybe by county with all points on top of that -- group by (county) -- and fill with color based on count  

#create scatterplot for the relationship between NP and bird sightings 

##run a lm on this^^^ relationship

#create boxplots for the relationship between NP Index and bird count 


##dive into the 5 most common birds analyzed above (same species)

###create boxplots for the relationship between NP index and each species count (differentiate by color)
```

